Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms

### 2018-09-23 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: fix numeration of slides
RSD: fixes in slides

\graphicspath{ {../pictures/}{img/} }
%set pages numeration
%\setbeamertemplate{footline}[frame number]
%\setlength\abovecaptionskip{-3pt}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber
}
\setbeamertemplate{headline}{}

@@

\begin{document}
\begin{frame}[noframenumbering,plain]
\titlepage
\end{frame}

@@

  \frametitle{Dimension reduction}
  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
  \begin{gather}
    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
    \min\{f(y): y\in D\}=\min\{f(y(x)): x\in [0,1]\} \nonumber
  \end{gather}
  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
  \begin{figure}[ht]
    \vspace*{-0.5cm}
    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
  \end{figure}
\end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Basic parallel optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^{k+j})\) and the function \(\varphi(y(x^{k+j}))\), \(j=\overline{1,p}\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
    compute objective \(f(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
    \end{displaymath}
    \item[Step 5.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Parallel optimization method with multiple evolvents}
  Using the multiple mapping allows solving initial problem by parallel solving the problems
  \[
  \min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
  \]
  on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
  separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
  particular processor are interpreted as the results of the trials in the rest problems (in the
  corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
  [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
  operations:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
    \item[Step 2.] Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
    blocking of the point \(y^k\) ).
    \item[Step 3.] Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
    trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
    \(x{}^{k_1} ,\dots,x{}^{k_s} \)
    \item[Step 4.] Inform the rest of processors about the trial results at the point \(y^k\).
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusions}
    %Already done:
    \begin{itemize}
      \item The smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
      \item The shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint. About 95\% of iterations are overhead to fight the auxiliary constraint.
      \item Rotated evolvents perform almost the same as the shifted ones but without overhead.
      \item Parallel optimization method shows up to 43x speedup on hard \(5d\) problems when using a set of rotated evolvents.
    \end{itemize}
    %Future work:
    %\begin{itemize}
    %  \item
    %\end{itemize}
\end{frame}

\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    \Large{Contacts:}
\vspace{0.5cm}

@@

    https://github.com/sovrasov
  \end{center}
\end{frame}

\end{document}

>>>>

\graphicspath{ {../pictures/}{img/} }

%set pages numeration
\newcommand\numbered{\setbeamertemplate{footline}{%
  \vspace{-10em}
   \raisebox{5pt}{\makebox[\paperwidth]{%
     \hfill\makebox[10pt]{%
       \usebeamerfont{footline}\usebeamercolor[fg]{footline}
       \insertframenumber}}}}}
\newcommand\unnumbered{\setbeamertemplate{footline}{}}

@@

\begin{document}
\numbered
{
\unnumbered
\begin{frame}[noframenumbering,plain]
\titlepage
\end{frame}
}

@@

  \frametitle{Dimension reduction}
  Peano-type curve \(y(x)\) allows to reduce the dimension of the original problem:
  \begin{gather}
    \lbrace y\in \mathbb{R}^N:-2^{-1}\leqslant y_i\leqslant 2^{-1},1\leqslant i\leqslant N\rbrace=\{y(x):0\leqslant x\leqslant 1\} \nonumber \\
    \min\{\varphi(y): y\in D\}=\min\{\varphi(y(x)): x\in [0,1]\} \nonumber
  \end{gather}
  \(y(x)\) is non-smooth function which continuously maps the segment \([0,1]\) to the hypercube \(D\).
  \begin{figure}[ht]
    \vspace*{-0.5cm}
    \subfloat{{\includegraphics[width=.35\textwidth]{peano3d.png} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{peano2d.png} }}
  \end{figure}
\end{center}
\end{frame}

@@

\begin{frame}
  \frametitle{Basic parallel optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 3.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
    compute objective \(\varphi(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
    \end{displaymath}
    \item[Step 4.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Parallel optimization method with multiple evolvents}
  Using the multiple mapping allows solving initial problem by parallel solving the problems
  \[
  \min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
  \]
  on a set of intervals $[0,1]$ by the basic method. Each one-dimensional problem is solved on a
  separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
  particular processor are interpreted as the results of the trials in the rest problems (in the
  corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
  [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
  operations:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
    \item[Step 2.] Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
    blocking of the point \(y^k\) ).
    \item[Step 3.] Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
    trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
    \(x{}^{k_1} ,\dots,x{}^{k_s} \)
    \item[Step 4.] Inform the rest of processors about the trial results at the point \(y^k\).
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusions}
    %Already done:
    \begin{itemize}
      \item The smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
      \item The shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint. About 95\% of iterations are overhead to fight the auxiliary constraint.
      \item Rotated evolvents perform almost the same as the shifted ones but without overhead.
      \item Parallel optimization method shows up to 43x speedup on hard \(5d\) problems when using a set of rotated evolvents.
    \end{itemize}
    %Future work:
    %\begin{itemize}
    %  \item
    %\end{itemize}
\end{frame}
{
\unnumbered
\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    \Large{Contacts:}
\vspace{0.5cm}

@@

    https://github.com/sovrasov
  \end{center}
\end{frame}
}
\end{document}
